### 2022-11-14 Data/code.Rmd
commit #add code

```{r}
data <- as.data.frame(data)

str(data)
```

# to split into traiing and test

```{r}
smp_size <- floor(.80 * nrow(data))
set.seed(123)
train_ind <- sample(seq_len(nrow(data)), size = smp_size)

training <- data[train_ind, ]
validation <- data[-train_ind, ]

summary(training)
```

# Model 1
```{r}
fit1 <- lm(GDP ~ TPEC + POP + TCMR + DPI + UNRATE + BSLP, data = training)

summary(fit1)
```

```{r}
library(car)

avPlots(fit1)
```

```{r}
res1 <- resid(fit1)
plot(fitted(fit1), res1)
abline(0,0)
```

```{r}
qqnorm(res1)
qqline(res1)
```

## Model 2 
```{r}
fit2 <- lm(GDP ~  + TCMR + DPI + UNRATE + BSLP, data = training)

summary(fit2)
```

```{r}
avPlots(fit2)
```
```{r}
res2 <- resid(fit2)
plot(fitted(fit2), res2)
abline(0,0)
```

```{r}
qqnorm(res2)
qqline(res2)
```

# Validation

```{r}
fit1v <- lm(GDP ~ TPEC + POP + TCMR + DPI + UNRATE + BSLP, data = validation)

summary(fit1v)
```

```{r}
fit2v <- lm(GDP ~  + TCMR + DPI + UNRATE + BSLP, data = validation)

summary(fit2v)
```

```{r}
AIC(fit1v, fit2v)
```